Update to: Combining global tree cover loss data with historical national forest cover maps to look at six decades of deforestation and forest fragmentation in Madagascar$^{\S}$

### 2018-10-11 update_2017/update_2017.Rnw
Update 2017

\documentclass[a4paper, 12pt, leqno, dvipsnames]{article} %leqno: numéro d'équation à gauche
\pagenumbering{arabic} % choose how to number the pages
\usepackage{a4wide}
\usepackage[utf8]{inputenc} % accents interprétés
\usepackage{graphicx}
%\usepackage{subfig}
%\usepackage[hmargin=2cm, vmargin = 2cm, noheadfoot]{geometry} %% Pour gérer le format des pages
%\usepackage{layout} %% Pour avoir la longueur des marges
\usepackage[round,sort]{natbib} %% Natbib is a popular style for formatting references.
%\usepackage{multibib}
%\newcites{secnm}{Bibliographie} 
%\usepackage{verbatim} % for multiline comments
\usepackage{amssymb} %symbole de maths
\usepackage{amsmath} %idem
%\usepackage{stmaryrd} %% Symbole flèche à l'envers
%\usepackage{amsfonts}
\usepackage[english]{babel} %% Les titres en anglais
\usepackage{array} %% Pour centrer verticalement le contenu d'un tableau, entre autres...
\setcounter{secnumdepth}{4} %% Profondeur des sections, subsections
\usepackage{setspace} %% Gère l'interligne: singlespacing, doublespacing
\usepackage{booktabs}
%\singlespacing
\usepackage{longtable}
\usepackage[table]{xcolor}
\setlength{\LTleft}{-5cm plus 1 fill}
\setlength{\LTright}{-5cm plus 1 fill}
\usepackage[colorlinks=true,citecolor=Blue,urlcolor=Maroon]{hyperref} %% Gère les hyperliens
\usepackage{lineno} %% Numérotation des lignes
%\linenumbers
\newcommand{\logit}{\text{logit}}
\newcommand{\bs}[1]{\boldsymbol{#1}}
\newcommand{\p}{\text{p}}
\newcommand{\R}{\textnormal{\sffamily\bfseries R}}
\newcommand{\pkg}[1]{{\fontseries{b}\selectfont #1}}

@@

\title{Combining global tree cover loss data with historical national
  forest cover maps to look at six decades of deforestation and forest
  fragmentation in Madagascar}
\author{}
\date{}

@@

\vspace{-1cm}
\begin{center}
  Ghislain Vieilledent$^{1,2,\star}$, Clovis Grinand$^{3}$,
  Fety A. Rakotomalala$^{3}$,\\
  Rija Ranaivosoa$^{4}$, Jean-Roger Rakotoarijaona$^{4}$,\\
  Thomas F. Allnutt$^{5,6}$, and Frédéric Achard$^{1}$\\
\end{center}

@@

{\small
  \begin{flushleft}  
    $[1]$ \textbf{Joint Research Center of the European Commission} -- Bio-economy Unit, I-21027 Ispra (VA), ITALY\\
    $[2]$ \textbf{CIRAD} -- UPR Forêts et Sociétés, F-34398 Montpellier, FRANCE\\
    $[3]$ \textbf{ETC Terra}, F-75020 Paris, FRANCE\\
    $[4]$ \textbf{Office National pour l'Environnement}, 101 Antananarivo, MADAGASCAR\\
    $[5]$ \textbf{Wildlife Conservation Society}, 101 Antananarivo, MADAGASCAR\\
    $[6]$ \textbf{GreenInfo Network}, Oakland, California, USA\\
    ~\\
    $[\star]$ \textbf{Corresponding author:}
    \textbackslash{E-mail}:~ghislain.vieilledent@cirad.fr
    \textbackslash{Phone}:~+39.033.278.3516\\
  \end{flushleft}}

@@

\doublespacing
\linenumbers

@@

  The island of Madagascar has a unique biodiversity, mainly located
  in the tropical forests of the island. This biodiversity is
  highly threatened by anthropogenic deforestation. Existing
  historical forest maps at national level are scattered and have
  substantial gaps which prevent an exhaustive assessment of long-term
  deforestation trends in Madagascar. In this study, we combined
  historical national forest cover maps (covering the period
  1953-2000) with a recent global annual tree cover loss dataset
  (2001-2014) to look at six decades of deforestation and forest
  fragmentation in Madagascar (from 1953 to 2014). We produced new
  forest cover maps at 30~m resolution for the year 1990 and annually
  from 2000 to 2014 over the full territory of Madagascar. We
  estimated that Madagascar has lost 44\% of its natural forest cover
  over the period 1953-2014 (including 37\% over the period
  1973-2014). Natural forests cover 8.9 Mha in 2014 (15\% of the
  national territory) and include 4.4 Mha (50\%) of moist forests, 2.6
  Mha (29\%) of dry forests, 1.7 Mha of spiny forests (19\%) and
  177,000 ha (2\%) of mangroves. Since 2005, the annual deforestation
  rate has progressively increased in Madagascar to reach 99,000 ha/yr
  during 2010-2014 (corresponding to a rate of 1.1\%/yr). Around half of
  the forest (46\%) is now located at less than 100~m from the forest
  edge. Our approach could be replicated to other developing countries with tropical forest.
  Accurate forest cover change maps can be used to assess the
  effectiveness of past and current conservation programs and
  implement new strategies for the future. In particular, forest maps
  and estimates can be used in the REDD+ framework which aims at
  ``Reducing Emissions from Deforestation and forest Degradation'' and
  for optimizing the current protected area network.

@@

Separated from the African continent and the Indian plate about 165
and 88 million years ago respectively \citep{Ali2008}, the flora and
fauna of Madagascar followed its own evolutionary path. Isolation
combined with a high number of micro-habitats \citep{Pearson2009} has
led to Madagascar's exceptional biodiversity both in term of number of
species and endemism in many taxonomic groups \citep{Crottini2012,
  Goodman2005}. Most of the biodiversity in Madagascar is concentrated
in the tropical forests of the island which can be divided into four
types: the moist forest in the East, the dry forest in the West, the
spiny forest in the South and the mangroves on the West coast
\citep{Vieilledent2016}. This unparalleled biodiversity is severely
threatened by deforestation \citep{Harper2007, Vieilledent2013}
associated with human activities such as slash-and-burn agriculture
and pasture \citep{Scales2011}. Tropical forests in Madagascar also
store a large amount of carbon (136 MgC.ha$^{-1}$ in the moist forest,
\citealp{Vieilledent2016}) and high rates of deforestation in
Madagascar (1.4--4.7~\%/yr, \citealp{Achard2002}) are responsible for
large CO$_2$ emissions in the atmosphere. Deforestation threatens
species survival by directly reducing their available habitat
\citep{Brooks2002, Tidd2001}. Forest fragmentation can also lead to
species extinction by isolating populations from each other and
creating forest patches too small to maintain viable populations
\citep{Saunders1991}. Fragmentation also increases forest edge where
ecological conditions (such as air temperature, light intensity and
air moisture) can be dramatically modified, with consequences on the
abundance and distribution of species \citep{Murcia1995,
  Broadbent2008, Gibson2013}. Forest fragmentation can also have
substantial effects on forest carbon storage capacity, as carbon
stocks are about 50\% lower at the forest edge than under a closed
canopy \citep{Brinck2017}. Moreover, forest carbon stocks vary
spatially due to climate or soil factors \citep{Saatchi2011,
  Vieilledent2016}. As a consequence, accurate and spatially explicit
maps of forest cover and forest cover change are necessary to monitor
biodiversity loss and carbon emissions from deforestation and forest
fragmentation, assess the efficiency of present conservation
strategies \citep{Eklund2016}, and implement new strategies for the
future \citep{Vieilledent2013, Vieilledent2016}. Simple time-series of
forest cover estimates, such as those provided by the FAO Forest
Resource Assessment report \citep{Keenan2015} are not sufficient.

Unfortunately, accurate and exhaustive forest cover maps are not
available for Madagascar after year 2000. \citet{Harper2007} produced
maps of forest cover and forest cover changes over Madagascar for the
years 1953, 1973, 1990 and 2000. The 1953 forest map is a vector map
derived from the visual interpretation of aerial photographs. Forest
maps for the years 1973, 1990, and 2000 were obtained from the
supervised classification of Landsat satellite images and can be used
to derive more accurate estimates of forest cover than those from the
FAO Forest Resource Assessment report. Nonetheless, maps provided by
\citet{Harper2007} are not exhaustive (due to the presence of clouds
in the satellite imagery), e.g.~11 244~km2 are mapped as unknown cover
type for the year 2000. Using a similar supervised classification
approach as in \citet{Harper2007}, more recent maps have been produced
for the periods 2000-2005-2010 by national institutions, with the
technical support of international environmental NGOs \citep{MEFT2009,
  ONE2013}. Another set of recent forest cover maps using an advanced
statistical tool for classification, the Random Forest classifier
\citep{Grinand2013, Rakotomalala2015}, was produced for the periods
2005-2010-2013 \citep{ONE2015}. However, these maps are either too old
to give recent estimates of deforestation \citep{MEFT2009, ONE2013},
include large areas of missing information due to images with high
percentage of cloud cover \citep{ONE2013}, or show large
mis-classification in specific areas, especially in the dry and spiny
forest domain, for which the spectral signal shows strong seasonal
variations due to the deciduousness of such forests (overall accuracy
is lower than 0.8 for the dry and spiny forests for the maps produced
by \citet{ONE2015}). Moreover, the production of such forest maps from
a supervised classification approach requires significant resources,
especially regarding the image selection step (required to minimize
cloud cover) and the training step (visual interpretation of a large
number of polygons needed to train the classification algorithm)
\citep{Rakotomalala2015}. Most of this work of image selection and
visual interpretation would need to be repeated to produce new forest
maps in the future using a similar approach.

Global forest or tree cover products have also been published recently
and can be tested at the national scale for
Madagascar. \citet{Kim2014} produced a global forest cover change map
from 1990 to 2000 (derived from Landsat imagery). This product was
updated to cover the period 1975-2005
(\url{http://glcf.umd.edu/data/landsatFCC/}) but forest cover maps
after 2005 were not produced. Moreover, the approach used in
\citet{Kim2014} did not accurately map the forests in the dry and
spiny ecosystems of Madagascar (see Fig. 8 in \citealt{Kim2014}).
\citet{Hansen2013} mapped tree cover percentage, annual tree cover
loss and gain from 2000 to 2012 at global scale at 30 m
resolution. This product has since been updated and is now available
up to the year 2014 \citep{Hansen2013}. To map forest cover from the
\citet{Hansen2013} product, a tree cover threshold must be selected
(that defines forest cover). Selecting such a threshold is not
straightforward as the accuracy of the global tree cover map strongly
varies between forest types, and is substantially lower for dry
forests than for moist forests \citep{Bastin2017}. Moreover, the
\citet{Hansen2013} product does not provide information on
land-use. In particular the global tree cover map does not separate
tree plantations such as oil palm or eucalyptus plantations from
natural forests \citep{Tropek2014}. Thus, the global tree cover map
from \citet{Hansen2013} cannot be used alone to produce a map of
forest cover \citep{Tyukavina2017}.

In this study, we present a simple approach which combines the
historical forest maps from \citet{Harper2007} and more recent global
products from \citet{Hansen2013} to derive annual wall-to-wall forest
cover change maps over the period 2000-2014 for Madagascar. We use the
forest cover map provided by \citet{Harper2007} for the year 2000
(defining the land-use) with the tree cover loss product provided by
\citet{Hansen2013} that we apply only inside forest areas identified
by \citet{Harper2007}. Similar to the approach of \citet{Harper2007},
we also assess trends in deforestation rates and forest fragmentation
from 1953 to 2014. We finally discuss the possibility to extend our
approach to other tropical countries or repeat it in the future for
Madagascar. We also discuss how our results could help assess the
effectiveness of past and current conservation strategies in
Madagascar, and implement new strategies in the future.

\section{Materials and Methods}
\label{materials-and-methods}

\subsection{Creation of new forest cover maps of Madagascar from
1953 to 2014}

Original 1990-2000 forest cover change map for Madagascar from
\citet{Harper2007} is a raster map at 28.5~m resolution. It was
derived from the supervised classification of Landsat TM (Thematic
Mapper) and ETM+ (Enhanced Thematic Mapper Plus) satellite images. For
our study, this map has been resampled at 30~m resolution using a
nearest-neighbor interpolation and reprojected in the WGS 84/UTM zone
38S projected coordinate system.

The 2000 Harper's forest map includes 208,000 ha of unclassified areas
due to the presence of clouds on satellite images. Unclassified areas
were mostly (88\%) present within the moist forest domain which
covered 4.17 Mha in 2000. To provide a label (forest or non-forest) to
these unclassified pixels, we used the 2000 tree cover percentage map
of \citet{Hansen2013} and selected a tree cover threshold of 75\% to
define the forest \citep{Achard2014, Aleman2017}. This threshold
allows to characterize properly the moist forest in Madagascar as 90\%
of the moist forest in 2000 in \citet{Harper2007} has a tree cover
greater than 75\% (Fig.~\ref{fig:tc_threshold}). For this step, the
Hansen's 2000 tree cover map was resampled on the same grid as the
original Harper's map at 30~m resolution using a bilinear
interpolation. We thus obtained a forest cover map for the year 2000
covering the full territory of Madagascar.

We then combined the forest cover map of the year 2000 with the annual
tree cover loss maps from 2001 to 2014 from \citet{Hansen2013} to
create annual forest cover maps from 2001 to 2014 at 30~m
resolution. To do so, Hansen's tree cover loss maps were resampled on
the same grid as the original Harper's map at 30~m resolution using a
nearest-neighbor interpolation. We also completed the Harper's forest
map of year 1990 by filling unclassified areas (due to the presence of
clouds on satellite images) using our forest cover map of year
2000. To do so, we assumed that if forest was present in 2000, the
pixel was also forested in 1990. Indeed, there is little evidence of
natural forest regeneration in Madagascar \citep{Grouzis2001,
  Harper2007}, especially over such a short period of time.  The
remaining unclassified pixels were limited to a relatively small total
area of about 8,000 ha. We labeled these residual pixels as
non-forest, as for the year 2000.

The 1973 forest cover map for Madagascar from \citet{Harper2007} is a
raster map at 57~m resolution derived from the supervised
classification of Landsat MSS (Multispectral Scanner System) satellite
images. We resampled this map at 30~m resolution using a
nearest-neighbor interpolation on the same grid as the forest cover
maps for years 1990 and 2000. We completed the Harper's forest map of
year 1973 by filling unclassified areas using our forest cover map of
the year 1990 assuming that if forest was present in 1990, it was also
present in 1973. Contrary to the year 1990, the remaining unclassified
pixels for year 1973 corresponded to a significant total area of 3.3
millions ha which was left as is.

The 1953 forest cover map from \citet{Harper2007} is a vector map
produced by scanning the paper map of \citet{Humbert1965} which was
derived from the visual interpretation of aerial photographs. We
reprojected the forest cover map of year 1953 in the WGS 84/UTM zone
38S projected coordinate system. Because of the methodology used to
derive the 1953 forest cover map, it was not possible to perfectly
aligned this map with the forest cover maps of later years which were
produced through digital processing of satellite imagery. As a
consequence, the 1953 cannot be merged with the map of later years to
identify precisely the location of the deforested areas. Nonetheless,
the 1953 forest cover map can be used to have a rough estimate of the
forest cover and forest fragmentation at this date.  To do so, the map
was rasterized at 30~m resolution on the same grid as the forest cover
maps for years 1973, 1990 and 2000.

Finally for all forest cover maps from 1973, isolated single
non-forest pixels (i.e.~fully surrounded by forest pixels) were
recategorized as forest pixels. Doing so, forest cover increased from
about 95,000~ha for year 1953 to about 600,000~ha for year 2010. This
allowed us to avoid counting very small scale events (\textless 0.1
ha, such as selective logging or wind-throw) as deforestation. It also
prevents us from underestimating forest cover and overestimating
forest fragmentation.

\subsection{Computing forest cover areas and deforestation rates}

From these new forest cover maps, we calculated the total forest cover
area for seven available years (1953-1973-1990-2000-2005-2010-2014),
and the annual deforested area and annual deforestation rate for the
corresponding six time periods between 1953 and 2014. The annual
deforestation rates were calculated using Eq.~\ref{eq:theta}
\citep{Puyravaud2003, Vieilledent2013}:

\begin{equation}
  \label{eq:theta}
  \theta = 100 \times [1-(1-(F_{t_2}-F_{t_1})/F_{t_1})^{(1/(t_2-t_1))}]
\end{equation}

In Eq.~\ref{eq:theta}, $\theta$ is the annual deforestation rate (in \%/yr),
$F_{t_2}$ and $F_{t_1}$ are the forest cover free of clouds at both
dates $t_2$ and $t_1$, and $t_2-t_1$ is the time-interval (in
years) between the two dates.

Because of the large unclassified area (3.3 millions ha) in 1973, the
annual deforestation areas and rates for the two periods 1953-1973 and
1973-1990 are only partial estimates computed on the basis of the
available forest extent. Area and rate estimates are produced at the
national scale and for the four forest types present in
Madagascar: moist forest in the East, dry forest in the West, spiny
forest in the South, and mangroves on the Western coast
(Fig.~\ref{fig:ecoregion}). To define the forest types, we used a
map from the MEFT (\emph{``Ministère de l'Environnement et des Forêts
  à Madagascar''}) with the boundaries of the four ecoregions in
Madagascar. Ecoregions were defined on the basis of climatic and
vegetation criteria using the climate classification by
\citet{Cornet1974} and the vegetation classification from the 1996
IEFN national forest inventory \citep{IEFN1996}. Because mangrove
forests are highly dynamic ecosystems that can expand or contract on
decadal scales depending on changes in environmental factors
\citep{Armitage2015}, a fixed delimitation of the mangrove ecoregion
on six decades might not be fully appropriate. As a consequence, our
estimates of the forest cover and deforestation rates for mangroves in
Madagascar must be considered with this limitation.

\subsection{Comparing our forest cover and deforestation rate
estimates with previous studies}

We compared our estimates of forest cover and deforestation rates with
estimates from the three existing studies at the national scale for
Madagascar: (i) \citet{Harper2007}, (ii) \citet{MEFT2009} and (iii)
\citet{ONE2015}. \citet{Harper2007} provides forest cover and
deforestation estimates for the periods
c. 1953-c. 1973-1990-2000. \citet{MEFT2009} provides estimates for the
periods 1990-2000-2005 and \citet{ONE2015} provides estimates for the
periods 2005-2010-2013. To compare our forest cover and deforestation
estimates over the same time periods, we consider an additional
time-period in our study (2010-2013) by creating an extra forest cover
map for the year 2013. We computed the Pearson's correlation
coefficient and the root mean square error (RMSE) between our
forest cover estimates and forest cover estimates from previous
studies for all the dates and forest types (including also the total
forest cover estimates). For previous studies, the computation of
annual deforestation rates (in \%/yr) is not always detailed and might
slightly differ from one study to another
\citep[see][]{Puyravaud2003}. \citet{Harper2007} also provide total
deforested areas for the two periods 1973-1990 and 1990-2000. We
converted these values into annual deforested area estimates. When
annual deforested areas were not reported (for 1953-1973 in
\citet{Harper2007} and in \citet{MEFT2009} and \citet{ONE2015}), we
computed them from the forest cover estimates in each study. These
estimates cannot be corrected from the potential bias due to the
presence of residual clouds. Forest cover and deforestation rates were
then compared between all studies for the whole of Madagascar and the
four ecoregions. The same ecoregion boundaries as in our study were
used in \citet{ONE2015} but this was not the case for
\citet{Harper2007} and \citet{MEFT2009}, which can explain a part of
the differences between the estimates.

\subsection{Fragmentation}

We also conducted an analysis of changes in forest fragmentation for
the years 1953, 1973, 1990, 2000, 2005, 2010 and 2014 at 30~m
resolution. We used a moving window of $51 \times 51$ pixels
(corresponding to an area of about 2.34~km$^2$) centered on each
forest pixel to compute the percentage of forest pixels in the
neighborhood. We used this percentage as an indication of the forest
fragmentation \citep{Riitters2012, Vogt2017}. The size of the moving
windows was based on a compromise: a sufficiently high number of cells
(here 2601) had to be considered to be able to compute a percentage
and a reasonably low number of cells had to be chosen to have a local
estimate of the fragmentation. Water bodies were not masked when
computing the percentage of forest pixels, meaning that forest located
near a water body was considered as fragmented. Computations were done
using the function \texttt{r.neighbors} of the GRASS GIS software
\citep{Neteler2008}. Using the density of forest in the neighborhood,
we defined five forest fragmentation classes: 0-20\% (highly
fragmented), 21-40\%, 41-60\%, 61-80\% and 81-100\% (lowly
fragmented). We reported the percentage of forest falling in each
fragmentation class for the six years and analyzed the dynamics of
fragmentation over the six decades.

We also computed the distance to forest edge for all forest
pixels for the years 1953, 1973, 1990, 2000, 2005, 2010 and 2014. For
that, we used the function \texttt{gdal\_proximity.py} of the GDAL
library (\url{http://www.gdal.org/}). We computed the mean and 90\%
quantiles (5\% and 95\%) of the distance to forest edge and looked at
the variation of these values over time. Previous studies have shown
that forest micro-habitats were mainly altered within the first 100~m
of the forest edge \citep{Brinck2017, Murcia1995,
  Broadbent2008}. Consequently, we also estimated the percentage of
forest within the first 100~m of the forest edge for each year and
looked at the variation of this percentage over the six decades.

\section{Results}
\label{results}

\subsection{Forest cover change and deforestation rates}

Natural forests in Madagascar covered 16.0 Mha in 1953, about 27\% of
the national territory of 587,041 km2. In 2014, the forest cover
dropped to 8.9 Mha, corresponding to about 15\% of the national
territory (Fig.~\ref{fig:fcc} and
Tab.~\ref{tab:forest_cover}). Madagascar has lost 44\% of its natural
forest between 1953 and 2014, including 37\% between 1973 and 2014
(Fig.~\ref{fig:fcc} and Tab.~\ref{tab:forest_cover}). In 2014 the
remaining 8.9 Mha of natural forest were distributed as follow: 4.4
Mha of moist forest (50\% of total forest cover), 2.6 Mha of dry
forest (29\%), 1.7 Mha of spiny forest (19\%) and 0.18 Mha (2\%) of
mangrove forest (Fig.~\ref{fig:ecoregion} and
Tab.~\ref{tab:comp_forest}).

The forest cover change map produced on the period 1953-2014
(Fig.~\ref{fig:fcc}) allows to identify hot-spots of
deforestation. Among the many recent hot-spots of deforestation
visible on the map for the period 2000-2014 , one is located at the
south of the CAZ (\emph{``Corridor Ankeniheny Zahamena''}) protected
area, in the moist forest at the east of Madagascar (see eastern zoom
in Fig.~\ref{fig:fcc}). Another major hot-spot of deforestation is
located around the Ranobe-PK32 new protected area, in the dry forest
at the south-west of Madagascar (see western zoom in
Fig.~\ref{fig:fcc}).

Regarding the deforestation trend, we observed a progressive decrease
of the deforestation rate after 1990 from 205,000 ha/yr (1.6\%/yr)
over the period 1973-1990 to 42,000 ha/yr (0.4\%/yr) over the period
2000-2005 (Tab.~\ref{tab:forest_cover}). Then from 2005, the
deforestation rate has progressively increased and has more than
doubled over the period 2010-2014 (99,000 ha/yr, 1.1\%/yr) compared to
2000-2005 (Tab.~\ref{tab:forest_cover}). The deforestation trend,
characterized by a progressive decrease of the deforestation rate over
the period 1990-2005 and a progressive increase of the deforestation
after 2005, is valid for all four forest types except the spiny forest
(Tab.~\ref{tab:comp_defor}). For the spiny forest, the deforestation
rate during the period 2010-2013 was lower than on the period
2005-2010 (Tab.~\ref{tab:comp_defor}).

\subsection{Comparison with previous forest cover change studies
in Madagascar}

Forest cover maps provided by previous studies over Madagascar were
not exhaustive (unclassified areas) due to the presence of clouds on
satellite images used to produce such maps. In \citet{Harper2007}, the
maps of years 1990 and 2000 include 0.5 and 1.12 Mha of unknown cover
type respectively. Proportions of unclassified areas are not reported
in the two other existing studies at the national level by
\citet{MEFT2009} and \citet{ONE2015}. With our approach, we produced
wall-to-wall forest cover change maps from 1990 to 2014 for the full
territory of Madagascar (Fig.~\ref{fig:fcc}). This allowed us to
produce more robust estimates of forest cover and deforestation rates
over this period (Tab.~\ref{tab:forest_cover}). Our forest cover
estimates over the period 1953-2013 (considering forest cover
estimates at national level and by ecoregions for all the available
dates) were well correlated (Pearson's correlation coefficient $=$
0.99) to estimates from the three previous studies
(Tab.~\ref{tab:comp_forest}) with a RMSE of 300,000 ha (6\% of the
mean forest cover of 4.8 Mha when considering all dates and forest
types together). These small differences can be partly attributed to
differences in ecoregion boundaries. Despite significant differences
in deforestation estimates (Tab.~\ref{tab:comp_defor}), a similar
deforestation trend was observed across studies with a decrease of
deforestation rates over the period 1990-2005, followed by a
progressive increase of the deforestation after 2005.

\subsection{Variation of forest fragmentation over time}

Forest fragmentation has progressively increased since 1953 in
Madagascar. We observed a continuous decrease of the mean distance to
forest edge from 1953 to 2014 in Madagascar. The mean distance to
forest edge has decreased to about 300~m in 2014 while it was of about
1.5~km in 1973 (Fig.~\ref{fig:dist_edge}). Moreover, a large
proportion (73\%) of the forest was located at a distance greater than
100~m in 1973, while almost half of the forest (46\%) is at a distance
lower than 100~m from forest edge in 2014
(Fig.~\ref{fig:dist_edge}). The percentage of lowly fragmented forest
in Madagascar has continuously decreased since 1953. The percentage of
forest belonging to the lowly fragmented class has fallen from 57\% in
1973 to 44\% in 2014. In 2014, 22\% of the forest belonged to the two
highest fragmented forest classes (less than 40\% of forest cover in
the neighborhood) while only 15\% of the forest belonged to these two
fragmentation classes in 1973 (Tab.~\ref{tab:frag}).

\section{Discussion}
\label{discussion}

\subsection{Advantages of combining recent global annual tree cover loss
  data with historical national forest cover maps}

In this study, we combined recent (2001-2014) global annual tree cover
loss data \citep{Hansen2013} with historical (1953-2000) national
forest cover maps \citep{Harper2007} to look at six decades
(1953-2014) of deforestation and forest fragmentation in
Madagascar. We produced annual forest cover maps at 30~m resolution
covering Madagascar for the period 2000 to 2014. Our study extends the
forest cover monitoring on a six decades period (from 1953 to 2014)
while harmonizing the data from previous studies \citep{Harper2007,
  MEFT2009, ONE2015}. We propose a generic approach to solve the
problem of forest definition which is needed to transform the 2000
global tree cover dataset from \citet{Hansen2013} into a
forest/non-forest map \citep{Tropek2014}. We propose the use of an
historical national forest cover map, based on a national forest
definition, as a forest cover mask. This approach could be easily
extended to other tropical regions or countries for which an accurate
forest cover map is available at any date within the period 2000-2014
(but preferably at the beginning of the period to profit from the full
record of tree cover loss and derive long-term estimates of
deforestation). For example, forest cover maps are available at 20~m
resolution for Cameroon and Central African Republic for years 2000
and 2010 \citep{Gross2017}. When high resolution forest cover maps are
not available, coarser resolution forest cover maps (leading to
coarser deforestation estimates) could be extracted from global land
cover products such as GLC2000 at 1~km resolution
\citep{Bartholome2005} or CCI Land Cover at 300~m resolution
\citep{Li2018}.  Moreover, this approach could be repeated in the
future with the release of updated tree cover loss data. We have made
the \R{}/GRASS code used for this study freely available in a GitHub
repository (see Data availability statement) to facilitate application
to other study areas or repeat the analysis in the future for
Madagascar.

The accuracy of the derived forest cover change maps depends directly
on the accuracy of the historical forest cover maps and the tree cover
loss dataset. Using visual-interpretation of aerial images in 342
areas distributed among all forest types, \citet{Harper2007} estimated
an overall 89.5\% accuracy in identifying forest/non-forest classes
for the year 2000. The accuracy assessment of the tree cover loss
dataset for the tropical biome reported 13\% of false positives and
16.9\% of false negatives (see Tab. S5 in \citealt{Hansen2013}). These
numbers rise at 20.7\% and 20.6\% respectively for the subtropical
biome. In the subtropical biome, the lower density tree cover canopy
makes it difficult to detect change from tree cover to bare
ground. For six countries in Central Africa, with a majority of moist
dense forest, \citet{Verhegghen2016} have compared deforestation
estimates derived from the global tree cover loss dataset
\citep{Hansen2013} with results derived from semi-automated supervised
classification of Landsat satellite images \citep{Achard2014} and they
found a good agreement between the two sets of estimates. Therefore,
our forest cover change maps after 2000 might be more accurate for the
dense moist forest than for the dry and spiny forest. In another study
assessing the accuracy of the tree cover loss product across the
tropics \citep{Tyukavina2015}, authors reported 4\% of false positives
and 48\% of false negatives in Sub-Saharan Africa. They showed that
85\% of missing loss occurred on the edges of other loss patches. This
means that tree cover loss might be underestimated in Sub-Saharan
Africa, probably due to the prevalence of small-scale disturbance
which is hard to map at 30~m, but that areas of large-scale
deforestation are well identified and spatial variability of the
deforestation is well represented. A proper accuracy assessment of our
forest cover change maps should be performed to better estimate the
uncertainty surrounding our forest cover change estimates in
Madagascar from year 2000 \citep{Olofsson2013,Olofsson2014}. Despite
this limitation, we have shown that the deforestation trend we
observed for Madagascar, with a doubling deforestation on the period
2010-2014 compared to 2000-2005, was consistent with the other studies
at the national scale \citep{ONE2015, MEFT2009}.

Consistent with \citet{Harper2007}, we did not consider potential
forest regrowth in Madagascar (although \citet{Hansen2013} provided a
tree cover gains layer for the period 2001-2013) for several
reasons. First, the tree gain layer of \citet{Hansen2013} includes and
catches more easily tree plantations than natural forest regrowth
\citep{Tropek2014}. Second, there is little evidence of natural forest
regeneration in Madagascar \citep{Grouzis2001, Harper2007}. This can
be explained by several ecological processes following burning
practice such as soil erosion \citep{Grinand2017} and reduced seed
bank due to fire and soil loss \citep{Grouzis2001}. Moreover, in areas
where forest regeneration is ecologically possible, young forest
regrowth are more easily re-burnt for agriculture and pasture. Third,
young secondary forests provide more limited ecosystem services
compared to old-growth natural forests in terms of biodiversity and
carbon storage \citep{Martin2013}.

\subsection{Natural forest cover change in Madagascar from 1953 to 2014}

>>>>

\documentclass[a4paper, 12pt, leqno, dvipsnames]{article} %leqno: numéro d'équation à gauche
\pagenumbering{arabic} % choose how to number the pages
\usepackage{a4wide}
\usepackage[utf8]{inputenc} % accents interprétés
\usepackage{graphicx}
%\usepackage{subfig}
%\usepackage[hmargin=2cm, vmargin = 2cm, noheadfoot]{geometry} %% Pour gérer le format des pages
%\usepackage{layout} %% Pour avoir la longueur des marges
\usepackage[round,sort]{natbib} %% Natbib is a popular style for formatting references.
%\usepackage{multibib}
%\newcites{secnm}{Bibliographie} 
%\usepackage{verbatim} % for multiline comments
\usepackage{amssymb} %symbole de maths
\usepackage{amsmath} %idem
%\usepackage{stmaryrd} %% Symbole flèche à l'envers
%\usepackage{amsfonts}
\usepackage[english]{babel} %% Les titres en anglais
\usepackage{array} %% Pour centrer verticalement le contenu d'un tableau, entre autres...
\setcounter{secnumdepth}{4} %% Profondeur des sections, subsections
\usepackage{setspace} %% Gère l'interligne: singlespacing, doublespacing
\usepackage{booktabs}
%\singlespacing
\usepackage{longtable}
\usepackage[table]{xcolor}
\setlength{\LTleft}{-5cm plus 1 fill}
\setlength{\LTright}{-5cm plus 1 fill}
\usepackage[colorlinks=true,citecolor=Blue,urlcolor=Maroon]{hyperref} %% Gère les hyperliens
%\usepackage{lineno} %% Numérotation des lignes
%\linenumbers
\newcommand{\logit}{\text{logit}}
\newcommand{\bs}[1]{\boldsymbol{#1}}
\newcommand{\p}{\text{p}}
\newcommand{\R}{\textnormal{\sffamily\bfseries R}}
\newcommand{\pkg}[1]{{\fontseries{b}\selectfont #1}}

@@

\title{Update to: Combining global tree cover loss data with historical national
  forest cover maps to look at six decades of deforestation and forest
  fragmentation in Madagascar$^{\S}$}
\author{}
\date{}

@@

\vspace{-1cm}
\begin{center}
  Ghislain Vieilledent$^{1,\star}$, Clovis Grinand$^{2}$,
  Fety A. Rakotomalala$^{2}$,\\
  Rija Ranaivosoa$^{3}$, Jean-Roger Rakotoarijaona$^{3}$,\\
  Thomas F. Allnutt$^{4,5}$, and Frédéric Achard$^{6}$\\
\end{center}

@@

{\small
\noindent$^{\S}$ This article was initially published as:

\noindent\textbf{Vieilledent G., C. Grinand, F. A. Rakotomalala, R. Ranaivosoa, J.-R. Rakotoarijaona, T. F. Allnutt, and F. Achard.} 2018. Combining global tree cover loss data with historical national forest-cover maps to look at six decades of deforestation and forest fragmentation in Madagascar. \emph{Biological Conservation.} \textbf{222}: 189-197. [doi: \href{https://doi.org/10.1016/j.biocon.2018.04.008}{10.1016/j.biocon.2018.04.008}].}

{\small
  \begin{flushleft}  
    $[1]$ \textbf{CIRAD} -- UPR Forêts et Sociétés, F-34398 Montpellier, FRANCE\\
    $[2]$ \textbf{Nitid\ae}, F-75020 Paris, FRANCE\\
    $[3]$ \textbf{Office National pour l'Environnement}, 101 Antananarivo, MADAGASCAR\\
    $[4]$ \textbf{Wildlife Conservation Society}, 101 Antananarivo, MADAGASCAR\\
    $[5]$ \textbf{GreenInfo Network}, Oakland, California, USA\\
    $[6]$ \textbf{Joint Research Center of the European Commission} -- Bio-economy Unit, I-21027 Ispra (VA), ITALY\\
    ~\\
    $[\star]$ \textbf{Corresponding author:}
    \textbackslash{E-mail}:~ghislain.vieilledent@cirad.fr
  \end{flushleft}}

@@

\doublespacing
%\linenumbers

@@

Blabla

@@

Blabla
